Predicting instabilities of a tuneable ring laser with an iterative map model

### 2020-12-03 Papers/Journal/Tuneable_Lasers.tex
My edits are done

\begin{abstract}
	Simple mathematical models have been unable to predict the conditions leading to instabilities in a tuneable ring laser. Here, we propose a nonlinear iterative map model for tuneable ring lasers. Solving a reduced nonlinear Schr\"odinger equation for each component in the laser cavity, we obtain an algebraic map for each component. Iterating the maps gives the total effect of one complete round trip. By neglecting the nonlinearity initially, we find a linearly chirped Gaussian to be the analytic fixed point solution, which we analyze asymptotically. We then numerically solve the full nonlinear model, allowing us to probe the underlying interplay of dispersion, modulation, and nonlinearity, as the pulse evolves over hundreds of round trips of the cavity. In the nonlinear case, we find the chirp saturates, and the envelope of the pulse becomes more rectangular in shape. Finally, for a nominal plane in the parameter space, we uncover a rich, sharp boundary separating the stable region and the unstable region where modulation instability degrades the pulse into an unsustainable state.
\end{abstract}

@@

Several effects arise within the cavity due to the interplay of dispersion, modulation, and the nonlinearity~\cite{bohun2015, coen1997, lapre2019, meng2020, oktem2010, shao2019, woodward2018}. The two effects of most interest for this paper are wave breaking, and modulation instability. Wave breaking causes the leading edge of a pulse to redshift, and the trailing edge to blueshift (the effect reverses in the anomalous dispersion regime) to the point a shock develops~\cite{anderson1992, rothenberg1989a, rothenberg1989b, tomlinson1984, tomlinson1985}. The frequency shift causes the pulse to become more rectangular in the frequency domain, with a linear chirp over most of the pulse, saturating at the leading and trailing edge. In optics, wave breaking manifests from self-phase modulation (SPM), which is a direct effect of the Kerr nonlinearity, causing the pulse to interfere with itself. New frequencies are then generated such that a rectangular profile is induced~\cite{agrawal2013, woodward2018}. Modulation instability---which typically arises in the anomalous dispersion regime---is the other effect that will be of interest. However, in the presence of multiple frequencies, modulation instability can emerge in the normal dispersion regime as well through cross-phase modulation and four-wave mixing~\cite{agrawal1987, agrawal2013, haelterman1992}. Modulation instability causes a pulse to degrade into much shorter, less intense pulses, and can cause a laser to lose coherence~\cite{agrawal1987, coen1997, haelterman1992}. Outside of optics, wave breaking and modulation instability occur in other areas with nonlinear waves and dispersive media such as plasma physics, heat propagation, transmission lines, and fluid dynamics~\cite{coen1997, rothenberg1989b}. Within a ring laser cavity wave breaking and modulation instability can become parasitic leading to an unstable and unsustainable pulse. Understanding the rich landscape of the parameter space, is important for determining design principles to guarantee the ring laser is stable and sustainable~\cite{bohun2015, burgoyneemail, finot2008, lapre2019, woodward2018}.

\begin{figure}[tbp]
	\centering
	\input{Cavity}
	\caption{Common cavity of a fibre ring laser~\cite{burgoyne2014, chung2017, lapre2019, shao2019}. The pump laser travels clockwise around the left-hand loop (red arrows) energizing the Er-doped fibre. The main laser travels clockwise around the right-hand loop (blue arrows) passing through each component, with most of the pulse exiting the cavity through the optical coupler.}
	\label{fig:cavity}
\end{figure}

The canonical equation in nonlinear optics is the nonlinear Schr\"odinger equation,
\begin{align}
	\pdiff{A}{z} &= - i \frac{\beta_2}{2}\pdiff[2]{A}{T} + i \gamma |A|^2 A.
	\label{eq:smallnlse}
\end{align}
Here, $A = A(T, z)$ is the complex pulse envelope, $T$ is co-moving time, $z$ is the propagation coordinate, and $\beta_2$ and $\gamma$ are the group velocity dispersion and Kerr nonlinearity, respectively. Typically, \eqref{eq:smallnlse} is generalized to incorporate terms more specific to laser physics. A gain term and loss term are added (and sometimes higher order dispersion terms), to give the generalized nonlinear Schr\"odinger equation (GNLSE)~\cite{agrawal2013, bohun2015, finot2008, peng2018, shtyrina2017, yarutkina2013},
	\begin{align}
	\pdiff{A}{z} &= - i \frac{\beta_2}{2}\pdiff[2]{A}{T} + i \gamma |A|^2 A + \frac{1}{2}g(A) A - \alpha A,
	\label{eq:nlse}
\end{align}
where $g(A)$ is an amplifying term owing to the gain fibre, and $\alpha$ is the loss due to scattering and absorption.

The GNLSE, \eqref{eq:nlse}, has applications in nonlinear optics and fibre optic communications, however, \eqref{eq:nlse} is regularly generalized even further to include a modulation term. The resulting equation is referred to as the master equation of mode-locking (see \cite{haus1975, haus1984, haus2000, tamura1996, usechak2005}), or sometimes the Ginzburg--Landau equation. Owing to the nonlinear term in \eqref{eq:nlse} and the often non-trivial form of $g(A)$, as well as the possible addition of a modulation term, no analytic solution is known to the master equation. See Haus~\cite{haus2000} for a comprehensive description and history of the theory of mode-locked lasers. While, the GNLSE well represents the dynamics in, say, a single optical fibre, \eqref{eq:nlse} is not entirely representative of the underlying physics within a laser cavity such as Fig. \ref{fig:cavity}. An assumption of \eqref{eq:nlse} is each process affects the pulse continuously within the cavity. Therefore, the GNLSE, and generalizations, can be thought of as an `average' model for lasers---this assumption breaks down if the pulse undergoes large changes during any one round trip. Often, this assumption does not hold because of the specialized, discrete components.

In this paper, we will build an iterative map model for the evolution of a laser pulse as the pulse travels though the cavity. In Section~\ref{sec:model}, we will briefly review previous iterative map models, then, we will derive our own iterative map model from \eqref{eq:nlse}. We will then non-dimensionalize the model and introduce four dimensionless parameters that govern the dynamics of the system. We will then examine the results of our model in Section~\ref{sec:results}---first solving the model analytically by assuming the absence of the nonlinearity, then numerically solving the full nonlinear model. We will identify a sharp boundary in the parameter space separating the region of stability, and the region where modulation instability degrades the pulse. Finally, the concluding thoughts and ideas for possible future work are given in Section~\ref{sec:conclusion}.

\section{Iterative Map Model}
\label{sec:model}
To resolve some of the issues with an `average' model, using separate GNLSEs for each component and region in the laser is common. Iteratively passing the pulse from one GNLSE to the next \cite{lapre2019, meng2020, oktem2010, woodward2018}. This method allows us to better encapsulate specific dynamics introduced by each component. The main drawback of this method is the extreme computational expense, since a nonlinear GNLSE must be numerically solved for every segment for each of the hundreds of round trips in a given simulation. We can alleviate this computational expense by noticing each effect is localized to a corresponding component in the cavity: virtually all of the dispersion happens within the CFBG~\cite{agrawal2002}, the pulse is amplified only within the Er-doped fibre, \emph{et cetera}. Therefore, we can drastically simplify each of the GNLSEs into a form we can solve analytically. Using analytic solutions instead of numerically solving a GNLSE reduces the computational cost, while maintaining the benefit of reflecting the specific geometry of a component-wise simulation.

Iterative map methods have been used for optics since 1955 when Cutler~\cite{cutler1955} analyzed a microwave regenerative pulse generator. Siegman and Kuizenga~\cite{kuizenga1970a, kuizenga1970b, kuizenga1970, siegman1969} adapted Cutler's method for both AM and FM mode-locked lasers in 1969, and Martinez \emph{et al.}~\cite{martinez1984, martinez1985} incorporated the nonlinearity while modelling passively mode-locked lasers. A linear iterative method was used for dispersion-tuned actively mode-locked lasers recently by Burgoyne \emph{et al.}~\cite{burgoyne2014}. Each of these models used a prescribed transfer function to describe the effect each component has on the pulse. While the transfer functions are chosen suitably to have the effect observed in the laboratory, the transfer functions are phenomenologically chosen, and not necessarily obtained directly from the underlying physics. Furthermore, these methods have not included the effect of every component, lacking the imperative interplay of dispersion, modulation, and nonlinearity---hindering their models to exhibit modulation instability.

We continue by deriving our own iterative map model by solving a reduced GNLSE for each component---except modulation, which we consider to be applied externally by an electro-optic modulator. Within the gain fibre we assume all other effects are negligible. Any dispersion, for example, in the gain fibre pales in comparison to the dispersion introduced by the CFBG. Thus, we neglect the dispersion (as well as loss and nonlinearity by the same argument) in the gain fibre, allowing us to simplify \eqref{eq:nlse} to
\begin{align}
	\pdiff{A}{z} &= \frac{1}{2}g(A)A.
	\label{eq:gainexample}
\end{align}
To proceed, we assume the gain has the form
\begin{align}
	g(A) &= \frac{g_0}{1 + E / \Es},
\end{align}
where $g_0$ is a small signal gain, $E$ is the energy of the pulse, defined as
\begin{align}
	E &= \int_{-\infty}^\infty |A|^2 \, \df T,
	\label{eq:energy}
\end{align}
and $\Es$ is the energy that the gain begins to saturate~\cite{haus1984, shtyrina2017, silfvast2004}, which depends on the power of the pump laser. Thus, \eqref{eq:gainexample} becomes
\begin{align}
	\pdiff{A}{z} &= \frac{g_0 A}{2 \left( 1 + E / \Es \right)}.
\end{align}
By assuming the energy entering the gain fibre is $E_g$ and the energy after travelling through the gain fibre of length $L_g$ is $E_{\text{out}}$, we find the energy after amplification is
\begin{align}
	\label{eq:engain}
	E_{\text{out}} &= \Es W \left( \frac{E_g}{\Es} \exp( E_g / \Es ) \exp( g_0 L_g ) \right),
\end{align}
where $W$ is the Lambert $W$ function. Rewriting \eqref{eq:engain} in terms of the pulse gives
\begin{align}
	G(A) &= \left( \frac{\Es}{E_g} W \left( \frac{E_g}{\Es} \exp( E_g / \Es ) \exp( g_0 L_g ) \right) \right)^{1/2} A
	\label{eq:gainform}
\end{align}
for the amplification of the pulse within the Er-doped fibre.

Repeating the same procedure for the loss, dispersion, and nonlinearity yields
\begin{align}
	L(A) &= (1 - R) \exp( - \alpha L_T )A, \\
	D(A) &= \FTi{\exp( i \omega^2 \beta_2 L_D/2 ) \FT{A}}, \\
	F(A) &= A \exp( i \gamma |A|^2 L_f ),
\end{align}
where $R$ is the reflectivity of the optical coupler, $L_T$ is the total length of the laser circuit, $L_D$ is the `length' of the dispersive medium, $L_f$ is the length of fibre between the amplifier and the optical coupler, and $\mathcal{F}$ denotes the Fourier transform. Finally, we consider the modulation. Here, we assume the pulse is modulated by an electro-optic modulator. For simplicity, the form is taken as the Gaussian
\begin{align}
	M(A) &= \exp( -T^2 / 2 T_M^2 ) A,
	\label{eq:modform}
\end{align}
where $T_M$ is the characteristic width of the modulation. This assumption is not restrictive. Calcaterra and Boldt~\cite{calcaterra2008a} have shown a sum of Gaussian functions can approximate any square integrable function. Therefore, our solution extends to any continuous, bounded modulation function. Additionally, using a saturable absorber in place of a modulator is common in practice. In this case, we can exchange \eqref{eq:modform} for~\cite{lapre2019, meng2020, oktem2010, woodward2018}
\begin{align}
	S(A) &= \left( 1 - \frac{q_0}{1 + |A|^2 / P_\text{sat}} \right) A,
\end{align}
where $q_0$ is the modulation depth, $|A|^2$ is the instantaneous power of the pulse, and $P_\text{sat}$ is the saturation power.

\subsection{Non-Dimensionalization}
Non-dimensionalization is a technique used in mathematical modelling to better understand the relative importance of different processes within a system. In essence, we scale all variables by typical values, based on the geometry and boundary conditions (see, for example,~\cite{howison2005}). Doing so `factors' the dimensional units out of the problem---leaving only dimensionless equations with dimensionless parameters. The relative magnitudes of the dimensionless parameters characterize the dominant dynamics---like how the Reynolds number can describe fluid flow. Each process in the laser can be better understood by re-scaling the time, energy, and amplitude by the convenient scalings:
\begin{align}
	T &= T_M \widetilde{T},& E &= \Es \widetilde{E},& A &= \left( \frac{\Es}{T_M} \right)^{1/2} \widetilde{A}.
\end{align}

Returning to the process maps \eqref{eq:gainform}--\eqref{eq:modform}, we find the non-dimensional mappings to be (after dropping the tildes)
\begin{align}
	G(A) &= \left(E_g^{-1} W \left( a E_g \textrm{e}^{E_g}\right) \right)^{1/2} A, & F(A) &= \exp( i b |A|^2) A, \nonumber \\
	D(A) &= \FTi{\exp( i s^2 \omega^2 ) \FT{A}}, & L(A) &= h A, \label{eq:effects} \\
	M(A) &= \exp( -T^2 / 2 ) A, \nonumber
\end{align}
with the four dimensionless parameters,
\begin{equation}
	\begin{aligned}
		a &= \exp( g_0 L_g ) \sim 8 \times 10^3, & 
		h &= (1 - R) \exp( - \alpha L_T) \sim 0.04, \\
		b &= \gamma L_f \frac{\Es}{T_M} \sim 1, & s &= \left( \frac{\beta_2 L_D}{2 T_M^2} \right)^{1/2} \sim 0.2,
		\label{eq:ndparam}
	\end{aligned}
\end{equation}
as defined by the values in Table~\ref{tab:values}. Here, $a$ is a measure of the strength of the gain fibre, $h$ is the total loss due to absorption and the optical coupler, $b$ is the ratio of the frequency shift because of the Kerr nonlinearity to the width of the pulse, and $s$ is the ratio of the effectiveness of the CFBG to the width of the pulse. Furthermore, in non-dimensional form, the modulation has no associated parameter, and each of the other processes in \eqref{eq:effects} have their own independent, non-dimensional parameter.

% Table of parameter values
\begin{table*}[tbp]
	\centering
	\caption{Range of parameter values (taken from~\cite{agrawal2013, burgoyne2014, burgoyneemail, tamura1996, usechak2005}).}
 	\label{tab:values}
 	\begin{tabular}{clll}
		\hline
		Parameter & Description & Value & Units \\
		\hline\noalign{\smallskip}
		$T_M$ & Characteristic modulation time & $15$--$150$ & ps \\
		$\beta_2 L_D$ & Dispersion provided by CFBG & $10$--$2000$ & ps$^2$ \\
		$\gamma$ & Kerr nonlinearity coefficient & $0.001$--$0.01$ & W$^{-1}$ m$^{-1}$ \\
		$L_f$ & Length of fibre between gain and optical coupler & $0.15$--$1$ & m \\
		$L_g$ & Length of the Er-doped gain fibre & $2$--$3$ & m \\
		$\alpha$ & Loss of fibre due to scattering and absorption & $10^{-4}$--$0.3$ & m$^{-1}$ \\
		$R$ & Reflectivity of optical coupler & $0.1$--$0.9$ & -- \\
		$\Es$ & Saturation energy of Er-doped gain fibre & $10^3$--$10^4$ & pJ \\
		$g_0$ & Small signal gain &  $1$--$10$ & m$^{-1}$ \\
		$L_T$ & Total length of fibre in the cavity & $10$--$100$ & m \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table*}

@@

\begin{figure}[tbp]
	\centering
	\includegraphics{Figures/Evo}
	\caption{Evolution of the envelope during one round trip of the cavity at equilibrium. The pulse decays because of the optical coupler, disperses owing to the CFBG, is modulated by the modulator, and finally, is amplified by the gain fibre (the envelope is unaltered by the nonlinearity). Note, the $z$ axis is not to scale.}
	\label{fig:cavityevo}
\end{figure}

Now that we have the algebraic effect of each section of the cavity, we are ready to take the composition of the maps to give the effect of one round trip of the cavity. Unlike the models described in the beginning of this section, the permutation of our components is important as the maps \eqref{eq:effects} do not necessarily commute with each other. The geometry of a particular laser will dictate the order of the process maps. We assume the construction of the laser is as in Fig.~\ref{fig:cavity}, and provide the following justifications. The effect of the nonlinearity will be maximal after the pulse passes through the gain fibre---after the gain is where the pulse is most energetic. Thus, we take the nonlinearity component to follow the gain. Similarly, we want the loss component to directly follow the nonlinearity in an attempt to minimize the effect the nonlinearity has on the pulse. All that remains is the dispersion and modulation components. For now, let us assume the CFBG precedes the modulator. Finally, we assume the loss component is our starting point since the loss corresponds to the optical coupler, and thus, allows us to more easily compare our results with experiments. More precisely, let
\begin{align}
	\mathcal{L}(A) = F(G(M(D(L(A)))))
	\label{eq:order}
\end{align}
denote one complete round trip of the cavity. Since a ring laser is cyclic in nature, we use the result of one round trip as the input pulse for the next round trip---restarting the process. We continually repeat this procedure until the envelope and chirp of the pulse reach equilibrium, or the pulse succumbs to modulation instability. The phase change is uninteresting to us, as \eqref{eq:effects} does not modify the phase in a meaningful way, and since the phase cannot easily be measured experimentally. An example of the evolution of the envelope during one round trip of the laser cavity can be found in Fig.~\ref{fig:cavityevo}.

@@

\begin{figure}[tbp]
	\centering
	\includegraphics{Figures/Sample_Gauss}
	\caption{Linearly chirped Gaussian pulse, \eqref{eq:A0}.}
	\label{fig:samplegauss}
\end{figure}

>>>>

\begin{abstract}
	Simple mathematical models have been unable to predict the conditions leading to instabilities in a tuneable ring laser. Here, we propose a nonlinear iterative map model for tuneable ring lasers. Solving a reduced nonlinear Schr\"odinger equation for each component in the laser cavity, we obtain an algebraic map for each component. Iterating through the maps gives the total effect of one round trip. By neglecting the nonlinearity initially, we find a linearly chirped Gaussian to be the analytic fixed point solution, which we analyze asymptotically. We then numerically solve the full nonlinear model, allowing us to probe the underlying interplay of dispersion, modulation, and nonlinearity, as the pulse evolves over hundreds of round trips of the cavity. In the nonlinear case, we find the chirp saturates, and the Fourier transform of the pulse becomes more rectangular in shape. Finally, for a nominal plane in the parameter space, we uncover a rich, sharp boundary separating the stable region and the unstable region where modulation instability degrades the pulse into an unsustainable state.
\end{abstract}

@@

Several effects arise within the cavity due to the interplay of dispersion, modulation, and the nonlinearity~\cite{bohun2015, coen1997, lapre2019, meng2020, oktem2010, shao2019, woodward2018}. The two effects of most interest for this paper are wave breaking, and modulation instability. Wave breaking causes the leading edge of a pulse to redshift, and the trailing edge to blueshift (the effect reverses in the anomalous dispersion regime) to the point a shock develops~\cite{anderson1992, rothenberg1989a, rothenberg1989b, tomlinson1984, tomlinson1985}. The frequency shift causes the pulse to become more rectangular in the frequency domain, with a linear chirp over most of the pulse, saturating at the leading and trailing edge. In optics, wave breaking manifests from self-phase modulation (SPM), which is a direct effect of the Kerr nonlinearity, causing the pulse to interfere with itself. New frequencies are then generated such that a rectangular profile is induced~\cite{agrawal2013, woodward2018}. Modulation instability---which typically arises in the anomalous dispersion regime---is the other effect that will be of interest. However, in the presence of multiple frequencies, modulation instability can emerge in the normal dispersion regime as well through cross-phase modulation and four-wave mixing~\cite{agrawal1987, agrawal2013, haelterman1992}. Modulation instability causes a pulse to degrade into much shorter, less intense pulses, and can cause a laser to lose coherence~\cite{agrawal1987, coen1997, haelterman1992}. Outside of optics, wave breaking and modulation instability occur in other areas with nonlinear waves and dispersive media such as plasma physics, heat propagation, transmission lines, and fluid dynamics~\cite{coen1997, rothenberg1989b}. Within a ring laser cavity, wave breaking and modulation instability can become parasitic leading to an unstable and unsustainable pulse. Understanding the rich landscape of the parameter space is important for determining design principles to guarantee the ring laser is stable and sustainable~\cite{bohun2015, burgoyneemail, finot2008, lapre2019, woodward2018}.

\begin{figure}[tbp]
	\centering
	\input{Cavity}
	\caption{Common cavity of a fibre ring laser~\cite{burgoyne2014, chung2017, lapre2019, shao2019}. The pump laser travels clockwise around the left-hand loop (red arrows) energizing the Er-doped fibre. The main laser travels clockwise around the right-hand loop (blue arrows), with most of the pulse exiting the cavity through the optical coupler.}
	\label{fig:cavity}
\end{figure}

The canonical equation in nonlinear optics is the nonlinear Schr\"odinger equation,
\begin{align}
	\pdiff{A}{z} &= - i \frac{\beta_2}{2}\pdiff[2]{A}{T} + i \gamma |A|^2 A.
	\label{eq:smallnlse}
\end{align}
Here, $A = A(T, z)$ is the complex pulse envelope, $T$ is co-moving time, $z$ is the propagation coordinate, and $\beta_2$ and $\gamma$ are the group velocity dispersion and Kerr nonlinearity, respectively. Typically, \eqref{eq:smallnlse} is generalized to incorporate terms more specific to laser physics. A gain term and a loss term are added (and sometimes higher order dispersion terms), to yield the generalized nonlinear Schr\"odinger equation (GNLSE)~\cite{agrawal2013, bohun2015, finot2008, peng2018, shtyrina2017, yarutkina2013},
	\begin{align}
	\pdiff{A}{z} &= - i \frac{\beta_2}{2}\pdiff[2]{A}{T} + i \gamma |A|^2 A + \frac{1}{2}g(A) A - \alpha A,
	\label{eq:nlse}
\end{align}
where $g(A)$ is an amplifying term owing to the gain fibre, and $\alpha$ is the loss due to scattering and absorption.

The GNLSE, \eqref{eq:nlse}, has applications in nonlinear optics and fibre optic communications, however, \eqref{eq:nlse} is regularly generalized even further to include a modulation term. The resulting equation is referred to as the master equation of mode-locking (see \cite{haus1975, haus1984, haus2000, tamura1996, usechak2005}), or sometimes the Ginzburg--Landau equation. Owing to the nonlinear term in \eqref{eq:nlse} and the often non-trivial form of $g(A)$, as well as the possible addition of a modulation term, no analytic solution is known to the master equation or to \eqref{eq:nlse}. See Haus~\cite{haus2000} for a comprehensive description and history of the theory of mode-locked lasers. While the GNLSE well represents the dynamics in, say, a single optical fibre, \eqref{eq:nlse} is not entirely representative of the underlying physics within a laser cavity such as Fig. \ref{fig:cavity}. An assumption of \eqref{eq:nlse} is each process affects the pulse continuously within the cavity. Therefore, the GNLSE, and generalizations, can be thought of as an `average' model for lasers---this assumption breaks down if the pulse undergoes large changes during any one round trip. Often, this assumption does not hold because of the specialized, discrete components.

In this paper, we will build an iterative map model for the evolution of a laser pulse as the pulse travels through the cavity. In Section~\ref{sec:model}, we will briefly review previous iterative map models, then, we will derive our own iterative map model from \eqref{eq:nlse}. We will then non-dimensionalize the model and introduce four dimensionless parameters that govern the dynamics of the system. We will then examine the results of our model in Section~\ref{sec:results}---first solving the model analytically by assuming the absence of the nonlinearity, then numerically solving the full nonlinear model. We will identify a sharp boundary in the parameter space separating the region of stability, and the region where modulation instability degrades the pulse. Finally, the concluding thoughts and ideas for possible future work are given in Section~\ref{sec:conclusion}.

\section{Iterative Map Model}
\label{sec:model}
To resolve some of the issues with an `average' model, using separate GNLSEs for each component and region in the laser is common. Iteratively passing the pulse from one GNLSE to the next \cite{lapre2019, meng2020, oktem2010, woodward2018}. This method allows for better encapsulation of specific dynamics introduced by each component. The main drawback of this method is the extreme computational expense since a nonlinear equation must be numerically solved for every segment for each of the hundreds of round trips in a given simulation. We can alleviate this computational expense by noticing each effect is localized to a corresponding component in the cavity: virtually all of the dispersion happens within the CFBG~\cite{agrawal2002}, the pulse is amplified only within the Er-doped fibre, \emph{et cetera}. Therefore, we can drastically simplify each of the GNLSEs into a form we can solve analytically. Using analytic maps instead of numerically solving a GNLSE reduces the computational cost, while maintaining the benefit of reflecting the specific geometry of a component-wise simulation.

Iterative map methods have been used for optics since 1955 when Cutler~\cite{cutler1955} analyzed a microwave regenerative pulse generator. Siegman and Kuizenga~\cite{kuizenga1970a, kuizenga1970b, kuizenga1970, siegman1969} adapted Cutler's method for both AM and FM mode-locked lasers in 1969, and Martinez \emph{et al.}~\cite{martinez1984, martinez1985} incorporated the nonlinearity while modelling passively mode-locked lasers. A linear iterative method was used for dispersion-tuned actively mode-locked lasers recently by Burgoyne \emph{et al.}~\cite{burgoyne2014}. Each of these models used a prescribed transfer function to describe the effect each component has on the pulse. While the transfer functions are chosen suitably to have the effect observed in the laboratory, the transfer functions are phenomenologically chosen, and not necessarily obtained directly from the underlying physics. Furthermore, these methods have not included the effect of every component, lacking the imperative interactions of dispersion, modulation, and nonlinearity---hindering their models to exhibit modulation instability.

We continue by deriving our own iterative map model by solving a reduced GNLSE for each component---except modulation, which we consider to be applied externally by an electro-optic modulator. Within the gain fibre, we assume all other effects are negligible. Any dispersion, for example, in the gain fibre pales in comparison to the dispersion introduced by the CFBG. Thus, we neglect the dispersion (as well as loss and nonlinearity by the same argument) in the gain fibre, allowing us to simplify \eqref{eq:nlse} to
\begin{align}
	\pdiff{A}{z} &= \frac{1}{2}g(A)A.
	\label{eq:gainexample}
\end{align}
To proceed, we assume the gain has the form
\begin{align}
	g(A) &= \frac{g_0}{1 + E / \Es},
\end{align}
where $g_0$ is a small signal gain, $E$ is the energy of the pulse, defined as
\begin{align}
	E &= \int_{-\infty}^\infty |A|^2 \, \df T,
	\label{eq:energy}
\end{align}
and $\Es$ is the energy that the gain begins to saturate~\cite{haus1984, shtyrina2017, silfvast2004}, which depends on the power of the pump laser. Thus, \eqref{eq:gainexample} becomes
\begin{align}
	\pdiff{A}{z} &= \frac{g_0 A}{2 \left( 1 + E / \Es \right)}.
\end{align}
By assuming the energy entering the gain fibre is $E_g$ and the energy after travelling through the gain fibre of length $L_g$ is $E_{\text{out}}$, we find the energy after amplification is
\begin{align}
	\label{eq:engain}
	E_{\text{out}} &= \Es W \left( \frac{E_g}{\Es} \exp( E_g / \Es ) \exp( g_0 L_g ) \right),
\end{align}
where $W$ is the Lambert $W$ function. Rewriting \eqref{eq:engain} in terms of the pulse gives
\begin{align}
	G(A) &= \left( \frac{\Es}{E_g} W \left( \frac{E_g}{\Es} \exp( E_g / \Es ) \exp( g_0 L_g ) \right) \right)^{1/2} A
	\label{eq:gainform}
\end{align}
for the amplification of the pulse after the Er-doped fibre.

Repeating the same procedure for the loss, dispersion, and nonlinearity yields
\begin{align}
	L(A) &= (1 - R) \exp( - \alpha L_T )A, \\
	D(A) &= \FTi{\exp( i \omega^2 \beta_2 L_D/2 ) \FT{A}}, \\
	F(A) &= A \exp( i \gamma |A|^2 L_f ),
\end{align}
where $R$ is the reflectivity of the optical coupler, $L_T$ is the total length of fibre in the cavity, $L_D$ is the `length' of the dispersive medium, $L_f$ is the length of fibre between the amplifier and the optical coupler, and $\mathcal{F}$ denotes the Fourier transform.

Finally, we consider the modulation. Here, we assume the pulse is modulated by an electro-optic modulator. For simplicity, the form is taken as the Gaussian
\begin{align}
	M(A) &= \exp( -T^2 / 2 T_M^2 ) A,
	\label{eq:modform}
\end{align}
where $T_M$ is the characteristic width of the modulation. This assumption is not restrictive. Calcaterra and Boldt~\cite{calcaterra2008a} have shown a sum of Gaussian functions can approximate any square integrable function. Therefore, our solution extends to any continuous, bounded modulation function. Additionally, using a saturable absorber in place of a modulator is common in practice. In this case, we can exchange \eqref{eq:modform} for~\cite{lapre2019, meng2020, oktem2010, woodward2018}
\begin{align}
	S(A) &= \left( 1 - \frac{q_0}{1 + |A|^2 / P_\text{sat}} \right) A,
\end{align}
where $q_0$ is the modulation depth, $|A|^2$ is the instantaneous power of the pulse, and $P_\text{sat}$ is the saturation power.

\subsection{Non-Dimensionalization}
Non-dimensionalization is a technique used in mathematical modelling to better understand the relative importance of different processes within a system. In essence, we scale all variables by typical values, based on the geometry and boundary conditions (see, for example,~\cite{howison2005}). Doing so `factors' the dimensional units out of the problem---leaving only dimensionless equations with dimensionless parameters. The relative magnitudes of the dimensionless parameters characterize the dominant dynamics---like how the Reynolds number describes fluid flow. Each process in the laser can be better understood by re-scaling the time, energy, and amplitude by the convenient scalings:
\begin{align}
	T &= T_M \widetilde{T},& E &= \Es \widetilde{E},& A &= \left( \frac{\Es}{T_M} \right)^{1/2} \widetilde{A}.
\end{align}

Returning to the process maps \eqref{eq:gainform}--\eqref{eq:modform}, we find the non-dimensional mappings to be (after dropping the tildes)
\begin{align}
	G(A) &= \left(E_g^{-1} W \left( a E_g \textrm{e}^{E_g}\right) \right)^{1/2} A, & F(A) &= \exp( i b |A|^2) A, \nonumber \\
	D(A) &= \FTi{\exp( i s^2 \omega^2 ) \FT{A}}, & L(A) &= h A, \label{eq:effects} \\
	M(A) &= \exp( -T^2 / 2 ) A, \nonumber
\end{align}
with the four dimensionless parameters,
\begin{equation}
	\begin{aligned}
		a &= \exp( g_0 L_g ) \sim 8 \times 10^3, & 
		h &= (1 - R) \exp( - \alpha L_T) \sim 0.04, \\
		b &= \gamma L_f \frac{\Es}{T_M} \sim 1, & s &= \left( \frac{\beta_2 L_D}{2 T_M^2} \right)^{1/2} \sim 0.2,
		\label{eq:ndparam}
	\end{aligned}
\end{equation}
as defined by the values in Table~\ref{tab:values}. Here, $a$ is a measure of the strength of the gain fibre, $h$ is the total loss due to absorption and the optical coupler, $b$ is the frequency shift because of the Kerr nonlinearity, and $s$ is the ratio of the effectiveness of the CFBG to the width of the pulse. Furthermore, in non-dimensional form, the modulation has no associated parameter, and each of the other processes in \eqref{eq:effects} have their own independent, non-dimensional parameter.

% Table of parameter values
\begin{table*}[tbp]
	\centering
	\caption{Range of parameter values (taken from~\cite{agrawal2013, burgoyne2014, burgoyneemail, tamura1996, usechak2005}).}
 	\label{tab:values}
 	\begin{tabular}{clll}
		\hline
		Parameter & Description & Value & Units \\
		\hline\noalign{\smallskip}
		$T_M$ & Characteristic modulation time & $15$--$150$ & ps \\
		$\beta_2 L_D$ & Dispersion provided by the CFBG & $10$--$2000$ & ps$^2$ \\
		$\gamma$ & Kerr nonlinearity coefficient & $0.001$--$0.01$ & W$^{-1}$ m$^{-1}$ \\
		$L_f$ & Length of fibre between gain and optical coupler & $0.15$--$1$ & m \\
		$L_g$ & Length of the Er-doped gain fibre & $2$--$3$ & m \\
		$\alpha$ & Loss of fibre due to scattering and absorption & $10^{-4}$--$0.3$ & m$^{-1}$ \\
		$R$ & Reflectivity of optical coupler & $0.1$--$0.9$ & -- \\
		$\Es$ & Saturation energy of Er-doped gain fibre & $10^3$--$10^4$ & pJ \\
		$g_0$ & Small signal gain &  $1$--$10$ & m$^{-1}$ \\
		$L_T$ & Total length of fibre in the cavity & $10$--$100$ & m \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table*}

@@

\begin{figure}[tbp]
	\centering
	\includegraphics{Figures/Evo}
	\caption{Evolution of the envelope during one round trip of the cavity at equilibrium. The pulse decays because of the optical coupler, disperses because of the CFBG, is modulated by the modulator, and finally, is amplified by the gain fibre (the envelope is unaltered by the nonlinearity). Note, the $z$ axis is not to scale.}
	\label{fig:cavityevo}
\end{figure}

Now that we have the algebraic effect of each section of the cavity, we are ready to take the composition of the maps to give the effect of one round trip of the cavity. Unlike the models described in the beginning of this section, the permutation of our components is important as the maps in \eqref{eq:effects} do not necessarily commute with each other.

The geometry of a particular laser will dictate the order of the process maps. However, we assume the construction of the laser is as in Fig.~\ref{fig:cavity}, and provide the following justifications. The effect of the nonlinearity will be maximal after the pulse passes through the gain fibre---after the gain is where the pulse is most energetic. Thus, we take the nonlinearity component to follow the gain. Similarly, we want the loss component to directly follow the nonlinearity in an attempt to minimize the effect the nonlinearity has on the pulse. All that remains is the dispersion and modulation components. For now, let us assume the CFBG precedes the modulator. Finally, we assume the loss component is our starting point since the loss corresponds to the optical coupler, and thus, allows us to more easily compare our results with experiments. More precisely, let
\begin{align}
	\mathcal{L}(A) = F(G(M(D(L(A)))))
	\label{eq:order}
\end{align}
denote one complete round trip of the cavity. Since a ring laser is cyclic in nature, we use the result of one round trip as the input pulse for the next round trip---restarting the process. We continually repeat this procedure until the envelope and chirp of the pulse reach equilibrium, or the pulse succumbs to modulation instability. The phase change is uninteresting to us, as \eqref{eq:effects} does not modify the phase in a meaningful way, and since the phase cannot easily be measured experimentally. An example of the evolution of the envelope during one round trip of the laser cavity can be found in Fig.~\ref{fig:cavityevo}.

@@

\begin{figure}[tbp]
	\centering
	\includegraphics{Figures/Sample_Gauss}
	\caption{Linearly chirped Gaussian pulse, \eqref{eq:A0} $(\phi_0 = 0)$.}
	\label{fig:samplegauss}
\end{figure}

\begin{figure*}[tbp]
	\centering
	\begin{subfigure}{\columnwidth}
		\centering
		\includegraphics{Figures/Chirp}
		\caption{Chirp.}
		\label{fig:chirp}
	\end{subfigure} %
	\begin{subfigure}{\columnwidth}
		\centering
		\includegraphics{Figures/FT}
		\caption{Fourier transform.}
		\label{fig:ft}
	\end{subfigure}
	\caption{Chirp and Fourier transform at equilibrium (after 500 round trips) initiated by \eqref{eq:nlA0} for three values of $b$, the nonlinearity parameter ($a = 8 \times 10^3$, $h = 0.04$, $s = 0.2$, and $E_0 = 0.1$).}
	\label{fig:chirpft}
\end{figure*}
